---
title: "group_4_MA7419_CW"
author: "hads2, sba11, js1392..."
date: "2025-11-17"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Questions

```{r message = FALSE}
library(readr)
cO2_data <- read_csv("Data/co-emissions-per-capita.csv")
electricty_pd <- read_csv("Data/electricity-prod-source-stacked.csv")
electricity_gt <- read_csv("Data/per-capita-electricity-generation.csv")
co2_sector <- read_csv("Data/per-capita-co2-sector.csv")
co2_vs_renw <- read_csv("Data/co2-per-capita-vs-renewable-electricity.csv")
```

## Slide 1: Jagroop

Title: How does Italy's reduction of carbon emissions compare to its
region and USA and China?

## Slide 2: Rahul

Overview Objectives: The objective is to assess Italy's relative
performance using key metrics like total emissions change, per-capita
figures, and renewable energy adoption.

1.How does Italy compares to other countries in CO2 emissions? - 1 line chart
2. What is Italy's Per capita CO₂ emissions by sector?
- 3 different graphs - bar, line, area
3. Italys per capita emissions through heat and electricity - 1 line chart
4. 


## Slide 3:

Method: Shahil

## Slide 4: Sahadat

<<<<<<< HEAD
Q1) How does Italy compare to ... in CO2 emissions?: 
graph of CO2 emissions, 1990-2024- compare italy to UK, France, Germany, EU, US, China short bullet points discussing the results, link to pulic interest
```{r}

library(ggplot2)
library(paletteer)
library(readr)
library(dplyr)
library(forcats)

Data1 <- cO2_data |>
  filter(Year >= 1990, Entity %in% c("Italy", "France", "Germany", "United Kingdom", "European Union (27)", "China", "United States")) |> 
  rename(emissions_pc = "Annual CO₂ emissions (per capita)") |>
  mutate(Entity = fct_reorder2(Entity, Year, emissions_pc))

                             
ggplot(Data1, aes(x = Year, y = emissions_pc, color = Entity, group = Entity)) +
  geom_line(size = 1.2, alpha = 0.8) +
  
  geom_line(data = filter(Data1, Entity == "Italy"), 
            size = 1.2, color = "Blue") +
  geom_point(data = filter(Data1, Entity == "Italy"),
             size = 2, color = "Blue", fill = "Blue", shape = 21) +
  
  scale_fill_paletteer_d("ggthemes::Tableu_10")+

  labs(
    title = "Comparing Italy's annual CO2 Emissions per Capita (1990 - 2024)",
    y = "per Capita CO2 Emissions (Million Tonnes)",
    X = NULL,
    color = "Country"
  ) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    legend.position = "right"
  )

# Analysis 
#Italy consistently emits less CO₂ per capita than Germany and the wider European Union average, while remaining higher than France.The United States emits nearly three times Italy’s CO2 per capita, but China follows an inverse trajectory, with rapid industrialization reaching almost 9 tonnes today.


```

=======
Q1) How does Italy compare to ... in CO2 emissions?: graph of CO2
emissions, 1990-2024- compare italy to UK, France, Germany, EU, US,
China short bullet points discussing the results, link to pulic interest
>>>>>>> 873bc7a6cb42344fd980dd266b0a7b0b92a5e8a7

## Slide 5: Hermann

Q2) What is Italy's Per capita CO₂ emissions by sector?

```{r}

library(dplyr)
library(janitor)
library(tidyverse)
library(stringr)
library(ggplot2)
library(gganimate)
library(forcats)
library(viridis)
library(ggthemes)
library(paletteer)
co2_sector <- clean_names(co2_sector, case = "snake") 
clCo2 <- co2_sector |> select(-per_capita_fugitive_emissions_of_carbon_dioxide_from_energy_production)



CS <- clCo2  |> rename(
  "Buildings" = per_capita_carbon_dioxide_emissions_from_buildings,
  "electricity and heat" = per_capita_carbon_dioxide_emissions_from_electricity_and_heat,
  "industry" = per_capita_carbon_dioxide_emissions_from_industry,
  "bunker fuels" = per_capita_carbon_dioxide_emissions_from_bunker_fuels,
  "land use change and forestry" = per_capita_carbon_dioxide_emissions_from_land_use_change_and_forestry,
  "emissions from transport" = per_capita_carbon_dioxide_emissions_from_transport,
  "manufacturing and construction" = per_capita_carbon_dioxide_emissions_from_manufacturing_and_construction,
  "other fuel combustion" = per_capita_carbon_dioxide_emissions_from_other_fuel_combustion)

CS <- CS |> pivot_longer(cols = c("Buildings",
                                   "electricity and heat",
                                   "industry",
                                   "bunker fuels",
                                   "land use change and forestry",
                                   "emissions from transport",
                                   "manufacturing and construction",
                                   "other fuel combustion"),
                          names_to = "sector",
                          values_to = "emission") |> 
  mutate(emission= round_half_up(emission, -1)) |> 
  arrange(desc(emission))

CS2 <- CS |> filter(year == 2022, code == "ITA") 


CS2 |> 
  filter(!is.na(emission)) |> 
  mutate(sector = fct_reorder(sector, emission)) |> 
  ggplot( aes(x=sector, y=emission,fill= sector)) +
  geom_bar(stat="identity", 
           alpha=.8, 
           width=.6, 
           colour = "black", 
           lineend =  "round", 
           linejoin = "mitre") +
  geom_text(aes(label = emission), hjust = "left", nudge_y = 12) +
  guides(fill="none") +
  coord_flip() +
  scale_fill_paletteer_d("ggthemes::Tableau_10") +
  theme(plot.background = element_rect(fill = "white"))+
  theme(axis.text.x = element_text(size = 8, face = "bold"),
        axis.text.y = element_text(size = 8, face = "bold"))+
  theme_minimal(base_size = 10, base_family = "sans")+
  labs(title = "Italy's 2022 CO2 emissions (kg) per capita by sector",
     x = "Emissions (Kg)",
     y = "Emissions (Kg)")

CS3 <- CS |> filter(entity == "Italy")

# 
# ggplot(CS3, aes(x=sector, y=emission, fill=sector)) +
#   geom_bar(stat='identity') +
#   guides(fill="none") +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

- Top Emitters: Electricity/heat (1,790) and transport (1,740) account for ~50% of emissions. Energy use and mobility are key targets.

- Mid-Tier: Buildings (890) and manufacturing (650) show moderate impact, indicating Italy's less industrialized economy.

- Carbon Sink: Land use/forestry (-220) provides a carbon offset.




```{r}


latest <- CS3 |> filter(year == max(year)) |> arrange(desc(emission))
CS3$sector <- factor(CS3$sector, levels = latest$sector)

ani <- ggplot(CS3, aes(x = year, y = emission, colour = sector)) +

ggplot(CS3, aes(x = year, y = emission, colour = sector)) +

  geom_line()+
  geom_point() +
  labs(title = "Italy's per capita CO2 emissions by sector",
     x = "",
     y = "Emissions (kg per capita)",
     )+
  labs(subtitle = "1990–2022")+
  scale_colour_manual(values = c(
  "electricity and heat" = "brown",
  "emissions from transport" = "darkgreen",
  "Buildings" = "darkblue",
  "manufacturing and construction" = "red",
  "bunker fuels" = "purple",
  "industry" = "orange",
  "other fuel combustion" = "#4575b4",
  "land use change and forestry" = "chocolate"
))+
  theme_minimal(base_size = 9, base_family = "sans")+
  scale_x_continuous(breaks = seq(1990, 2022, 5)) +
  theme(plot.background = element_rect(fill = "white"))+
  theme(axis.text.x = element_text(size = 8, face = "bold"),
        axis.text.y = element_text(size = 8, face = "bold"), 
        axis.title.y = element_text(size = 8, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", color = "brown"),
        legend.title = element_text(family = "Roboto",
                                    face = "bold",
                                    size = 8),
        legend.text  = element_text(size = 8))+
  theme(legend.position = "right")+
  transition_reveal(year) 

animate(
  ani,
  fps = 25,
  duration = 12,
  end_pause = 220
)

```
:::

## version 2

```{r}
ggplot(CS3, aes(x = year, y = emission, colour = sector)) +
  geom_line()+
  geom_point() +
  labs(title = "Italy's per capita CO2 emissions by sector",
     x = "",
     y = "Emissions (kg per capita)",
     )+
  labs(subtitle = "1990–2022")+
  scale_colour_manual(values = c(
  "electricity and heat" = "brown",
  "emissions from transport" = "darkgreen",
  "Buildings" = "darkblue",
  "manufacturing and construction" = "red",
  "bunker fuels" = "purple",
  "industry" = "orange",
  "other fuel combustion" = "#4575b4",
  "land use change and forestry" = "chocolate"
))+
  theme_minimal(base_size = 9, base_family = "sans")+
  scale_x_continuous(breaks = seq(1990, 2022, 5)) +
  theme(plot.background = element_rect(fill = "white"))+
  theme(axis.text.x = element_text(size = 8, face = "bold"),
        axis.text.y = element_text(size = 8, face = "bold"), 
        axis.title.y = element_text(size = 8, face = "bold"),
        plot.title = element_text(size = 12, face = "bold", color = "brown"),
        legend.title = element_text(family = "Roboto",
                                    face = "bold",
                                    size = 8),
        legend.text  = element_text(size = 6))+
  theme(legend.position = "right")
```

## version 3

```{r}


CS3 |> 
  ggplot(aes(x = year, y = emission, fill = sector)) +
  geom_area(alpha = 0.75) +
  labs(
    title = "Italy's per capita CO2 emissions by sector",
    subtitle = "1990–2022 • Stacked contributions of major sectors",
    y = "Emissions (kg per capita)",
    x = "",
    fill = "Sector"
  ) +
  scale_fill_paletteer_d("MexBrewer::Huida") +
  scale_x_continuous(breaks = seq(1990, 2022, 5)) +
  theme_minimal(base_size = 10) +
  theme(plot.background = element_rect(fill = "white"))+
  theme(
    legend.position = "right",
    legend.title = element_text(size = 9, face = "bold"),
    legend.text  = element_text(size = 6),
    plot.title = element_text(size = 16, face = "bold", colour = "chocolate4"),
    plot.subtitle = element_text(size = 11),
    axis.title.y = element_text(size = 10),
    axis.text.x = element_text(size = 8, face = "bold"),
    axis.text.y = element_text(size = 8, face = "bold")
  )


```
## Slide 6: Rahul

line graph of electricity and heat production for Italy vs .... discuss
the reulst and link to public interest

```{r}
library(dplyr)
library(ggplot2)
elec_heat <- co2_sector |> 
  rename(country = entity, co2_from_heat_electricity = per_capita_carbon_dioxide_emissions_from_electricity_and_heat) |>
  filter(country %in% c("Italy", "European Union (27)", "France", "Germany", "China", "United States", "United Kingdom")) |>
  select(country,code,year,co2_from_heat_electricity)

max_elec_heat <- elec_heat |>
  filter(year == max(year))|>
  arrange(desc(co2_from_heat_electricity))

elec_heat$country <- factor(elec_heat$country, levels = max_elec_heat$country)

ggplot(elec_heat, aes(x = year, y = co2_from_heat_electricity, color = country))+
  geom_line(key_glyph = "timeseries",size=1.3, alpha = .5)+
  labs(title = "Per capita CO2 emissions from electricity and heat",
       subtitle = "Italy compared with UK, France, Germany, China and US", 
       x = "", 
       y= "Per capita CO2 emissions",
       colour = "Country")+
  theme_classic()+
  theme(plot.title = element_text(face = "bold"),
        axis.title = element_text(face = "bold"))+
  geom_line(key_glyph = "timeseries",
            data = elec_heat |> filter(country=="Italy"),
            aes(x = year, y = co2_from_heat_electricity),
            size = 2)
```


## Slide 7: SHahil maybe

Q3) what percentage of Italy's energy production come form renewable
sources?

Line graph - percentage of Italy's energy production for 1990-2022

```{r}
# Cleaned/filtered data  
engery_clean <- co2_vs_renw |> 
  clean_names() |> 
  select(
    entity,
    year,
    annual_co2_emissions_per_capita,
    renewables_percent_electricity
  ) |> 
  filter(
    entity %in% c("Italy","United Kingdom", "France","China", "United States", "Germany"),
    year >= 1990 & year <= 2022 
  ) |> 
  filter(!is.na(renewables_percent_electricity))

# Checking if all country have full coverage for the years
engery_clean |> 
  group_by(entity) |> 
  summarise(
    min_year = min(year),
    max_year = max(year),
    n = n()
  )
View(engery_clean)
```

```{r}
# Plotting  
ggplot(engery_clean, aes(
  x = year ,
  y = renewables_percent_electricity,
  colour = entity
)) +
  geom_line( 
    size = 1) +
  
  geom_point(size = 1.5, alpha = 1) +
  
  geom_line(
    data = engery_clean |> filter(entity == "Italy"),
    aes(x = year,
        y = renewables_percent_electricity),
    linewidth = 1.5,
  ) +
  geom_point(
    data = engery_clean |> filter(entity == "Italy"),
    aes(x = year,
        y = renewables_percent_electricity),
    size = 2.2
  ) +
  labs(
    title = "Renewable Electricity  (1990-2022)",
    x = "Year",
    y = "Renewable Share of Electricity",
    subtitle = "Italy compared with UK, France, Germany, China and US",
    colour = "Country",
    caption = "Data source: Global Carbon Budget (2025) and other sources"
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10)
)
```
- Italy's renewable electricity share has seen a steady growth from 1990 to 2022, with massive growth starting from 2008.

- Italy outperforms the US, France and China, while the UK and Germany show faster renewable growth peaks. 

- Italy's progress shows commitment towards cleaner electricity usage and their transition to achieve that goal. 

100 word:
This line graph showcases the percentage of electricity produced from renewable energy sources, and is compared with the UK, Germany, France, China and US from 1990 to 2022. Italy generation of renewable electricity has increased with significant acceleration beginning from 2008. Also, Italy outperforms the US, France and China but is behind Germany and UK who show faster growth. The trends suggest all countries are committed to transitioning to renewable electricity which the graph shows meaning Italy progress towards more sustainable energy is positive.


## Slide 8: Jagroop

Pie chart - percentage of Italy's energy production for 22 discuss the
results and how it links to the public interest.

discuss the results and how it links to the public interest.

## Slide 9: Sahadat

Q4) Italy's environmental policy campare to ....?

## Slide 10: Hermann

Conclusion
:::
