---
title: "group_4_MA7419_CW"
author: "hads2, sba11, js1392..."
date: "2025-11-17"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Questions

```{r message = FALSE}
library(readr)
emissions_pc <- read_csv("Data/co-emissions-per-capita.csv")
electricty_pd <- read_csv("Data/electricity-prod-source-stacked.csv")
electricity_gt <- read_csv("Data/per-capita-electricity-generation.csv")
co2_sector <- read_csv("Data/per-capita-co2-sector.csv")
co2_vs_renw <- read_csv("Data/co2-per-capita-vs-renewable-electricity.csv")
```

## Slide 1: Jagroop

Title: How does Italy's reduction of carbon emissions compare to its
region and USA and China?

## Slide 2: Rahul

Overview Objectives: The objective is to assess Italy's relative
performance using key metrics like total emissions change, per-capita
figures, and renewable energy adoption.

## Slide 3:

Method: Shahil

## Slide 4: Sahadat

Q1) How does Italy compare to ... in CO2 emissions?: 
graph of CO2 emissions, 1990-2024- compare italy to UK, France, Germany, EU, US,
China short bullet points discussing the results, link to pulic interest

## Slide 5: Hermann

Q2) What is Italy's Per capita CO₂ emissions by sector? draph:
graph of taly's Per capita CO₂ emissions by sector (bar chart) for 2022

```{r}

library(dplyr)
library(janitor)
library(tidyr)
library(ggplot2)
library(gganimate)
library(forcats)
co2_sector <- clean_names(co2_sector, case = "snake") 
clCo2 <- co2_sector |> select(-per_capita_fugitive_emissions_of_carbon_dioxide_from_energy_production)



CS <- clCo2  |> rename(
  "Buildings" = per_capita_carbon_dioxide_emissions_from_buildings,
  "electricity_and_heat" = per_capita_carbon_dioxide_emissions_from_electricity_and_heat,
  "industry" = per_capita_carbon_dioxide_emissions_from_industry,
  "bunker_fuels" = per_capita_carbon_dioxide_emissions_from_bunker_fuels,
  "land_use_change_and_forestry" = per_capita_carbon_dioxide_emissions_from_land_use_change_and_forestry,
  "emissions_from_transport" = per_capita_carbon_dioxide_emissions_from_transport,
  "manufacturing_and_construction" = per_capita_carbon_dioxide_emissions_from_manufacturing_and_construction,
  "other_fuel_combustion" = per_capita_carbon_dioxide_emissions_from_other_fuel_combustion)
# 
CS <- CS |> pivot_longer(cols = c(Buildings,
                                   electricity_and_heat,
                                   industry,
                                   bunker_fuels,
                                   land_use_change_and_forestry,
                                   emissions_from_transport,
                                   manufacturing_and_construction,
                                   other_fuel_combustion),
                          names_to = "sector",
                          values_to = "emission") |> 
  mutate(emission= round_half_up(emission, -1)) |> 
  arrange(desc(emission))

CS2 <- CS |> filter(year == 2022, code == "ITA") 



# ggplot(CS2, aes(x=sector, y=emission)) + 
#   geom_bar(stat = "identity") +
#   coord_flip()

CS2 %>%
  filter(!is.na(emission)) |> 
  mutate(sector = fct_reorder(sector, emission)) %>%
  ggplot( aes(x=sector, y=emission,fill= sector)) +
  geom_bar(stat="identity", alpha=.6, width=.4) +
  geom_text(aes(label = emission), hjust = "left") +
  guides(fill="none") +
  coord_flip() +
  xlab("") +
  theme_bw()

CS3 <- CS |> filter(entity == "Italy") 


ggplot(CS3, aes(x=sector, y=emission, fill=sector)) +
  geom_bar(stat='identity')

```

```{r}
ggplot(CS3, aes(x=sector, y=emission, fill=sector)) +
  geom_bar(stat='identity') +
  theme_bw() +

  transition_states(
    year,
    transition_length = 33,
    state_length = 1,
    wrap = FALSE
  ) +
  guides(fill="none") +
  coord_flip() +
   ggtitle('Co2 emissions in {closest_state}',
          subtitle = 'Frame {frame} of {nframes}')
  ease_aes('sine-in-out')
```


```{r}
blue <- CS3 |> 
  filter(!is.na(emission)) |> 
  mutate(sector = fct_reorder(sector, emission)) |> 
  ggplot(aes(x=sector, y=emission, fill=sector)) + 
  geom_bar(stat='identity') +
  geom_text(aes(label = sprintf("%d", round(emission))), hjust = "left") +
  theme_bw() +
  transition_states(year)+
  guides(fill="none") +
  coord_flip() +
   ggtitle('Co2 emissions in {closest_state}',
          subtitle = 'Frame {frame} of {nframes}') 
  ease_aes('linear-in')

animate(blue,
        duration = 30,
        fps = 10,
        width = 800,
        height = 500,
        start_pause = 10, 
        end_pause = 20)
```


```{r}
try <- CS3 |> 
  filter(!is.na(emission)) |> 
  mutate(sector = fct_reorder(sector, emission),
         rank = rank(emission)) |> 
  # ggplot(aes(x=sector, y=emission, fill=sector)) + 
  ggplot(aes(xmin = 0,
             xmax = emission,
             y = rank,
             ymin = rank - 20,
             ymax = rank + 20,
             fill = sector,
             label = round(rank,0)
             )) +
  geom_rect(alpha = 0.5) +
  geom_text(aes(x = emission, label = as.character(emission)), hjust = "left") +
  geom_text(aes(x = 0, label = sector), hjust = "left") +
  theme_bw() +
  # Adjust the x-axis scale limits
  scale_x_continuous(limits = c(0, 300)) +
  
  # Customize labels and titles
  labs(x = NULL, y = NULL, title = "Emissions by sector") +
  
  # Add a label indicating the month
  geom_label(aes(sprintf("%d", round(year))), 
             x = 250, y = 1, 
             fill = "white", col = "black",
             size = 10, 
             label.padding = unit(1, "lines")) +
  
  # Apply a classic theme
  theme_classic() +
  
  # Customize plot appearance
  theme(legend.position = "none",
        axis.line.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.x = element_blank(),
        title = element_text(size = 40, hjust = 0.5)) +
  
  # Create multiple subplots for each month
  facet_wrap(~year) +
  
  # Remove facet labels: this allows us to superlay the facets on top of each other
  # and then animate the facets
  facet_null() + 
  
  # Create a time-based animation based on 'month_anim'
  transition_time(year)

# Animate the ggplot object with specified settings
animate(try,
        duration = 30,
        fps = 10,
        width = 800,
        height = 500,
        start_pause = 10, 
        end_pause = 20)
```

```{r}
try2 <-   ggplot(CS3, aes(x = emission, y = reorder(sector, emission), fill = sector)) +
  geom_col(width = 0.8) +
  geom_text(aes(label = sprintf("%d", round(emission))), hjust = "left") +
  transition_time(year) +
  theme_classic() +
  theme(legend.position = "none")

animate(try2,
        duration = 30,
        fps = 10,
        width = 800,
        height = 500,
        start_pause = 10, 
        end_pause = 20)
```

## Slide 6: Rahul

line graph of electricity and heat production for Italy vs ....
discuss the reulst and link to public interest 

## Slide 7: SHahil maybe

Q3) what percentage of Italy's energy production come form renewable
sources?

Line graph - percentage of Italy's energy production for 1990-2022

## Slide 8: Jagroop
Pie chart - percentage of Italy's energy production for 22
discuss the results and how it links to the public interest.


discuss the results and how it links to the public interest.

## Slide 9: Sahadat

Q4) Italy's environmental policy campare to ....?

## Slide 10: Hermann

Conclusion
