---
title: "group_4_MA7419_CW"
author: "hads2, sba11, js1392..."
date: "2025-11-17"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Questions

```{r message = FALSE}
library(readr)
emissions_pc <- read_csv("Data/co-emissions-per-capita.csv")
electricty_pd <- read_csv("Data/electricity-prod-source-stacked.csv")
electricity_gt <- read_csv("Data/per-capita-electricity-generation.csv")
co2_sector <- read_csv("Data/per-capita-co2-sector.csv")
co2_vs_renw <- read_csv("Data/co2-per-capita-vs-renewable-electricity.csv")
```

## Slide 1: Jagroop

Title: How does Italy's reduction of carbon emissions compare to its
region and USA and China?

## Slide 2: Rahul

Overview Objectives: The objective is to assess Italy's relative
performance using key metrics like total emissions change, per-capita
figures, and renewable energy adoption.

## Slide 3:

Method: Shahil

## Slide 4: Sahadat

Q1) How does Italy compare to ... in CO2 emissions?: 
graph of CO2 emissions, 1990-2024- compare italy to UK, France, Germany, EU, US,
China short bullet points discussing the results, link to pulic interest

## Slide 5: Hermann

Q2) What is Italy's Per capita CO₂ emissions by sector? draph:
graph of taly's Per capita CO₂ emissions by sector (bar chart) for 2022

## Slide 6: Rahul

line graph of electricity and heat production for Italy vs ....
discuss the reulst and link to public interest 

## Slide 7: SHahil maybe

Q3) what percentage of Italy's energy production come form renewable
sources?

Line graph - percentage of Italy's energy production for 1990-2022

## Slide 8: Jagroop
Pie chart - percentage of Italy's energy production for 22
discuss the results and how it links to the public interest.


discuss the results and how it links to the public interest.

## Slide 9: Sahadat

Q4) Italy's environmental policy campare to ....?

## Slide 10: Hermann

Conclusion
