---
title: "How does Italy's reduction of carbon emissions compare to its region, USA and China?"
format: revealjs
incremental: true
logo: CopyOfuniversity-of-leicester-logo_2.png
css: custom.css
editor: visual
theme: simple
fontsize: 1.3em
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message = FALSE}
library(readr)
library(dplyr)
library(janitor)
library(tidyverse)
library(stringr)
library(ggplot2)
library(gganimate)
library(forcats)
library(viridis)
library(ggthemes)
library(paletteer)
library(palette)
library(forcats)
library(plotly)
library(tidyr)

cO2_data <- read_csv("Data/co-emissions-per-capita.csv")
electricty_pd <- read_csv("Data/electricity-prod-source-stacked.csv")
electricity_gt <- read_csv("Data/per-capita-electricity-generation.csv")
co2_sector <- read_csv("Data/per-capita-co2-sector.csv")
co2_vs_renw <- read_csv("Data/co2-per-capita-vs-renewable-electricity.csv")
```

## Slide 1: Rahul

Overview Objectives: The objective is to assess Italy's relative performance using key metrics like total emissions change , per-capita figures, and renewable energy adoption.

## Method

| Step | Process | Justification |
|------------------------|------------------------|------------------------|
| **Research Question** | Investigate Italy's reduction of carbon emission using key metrics like renewable electricity and emissions while comparing it. | This ensures performance is compared globally for environmental improvement which Italy made. |
| **Data Sources** | The datasets are from Our World in Data (OWID) which all visuals have been created from. | These sources are valid because OWID is reliable, public and accessible for all countries to be compared. |
| **Data Cleaning** | Column names are cleaned and renamed appropriately and the datasets are filtered to relevant countries and years while N/A values are removed for better comparison. | This improves consistency and accuracy increases meaning fairer comparison between all relevant countries. |
| **Visuals** | Implemented pie chart, line graphs and highlighted area chart based on the question while highlighting Italy. | The area chart and line chart both showcase the trends over time between the variables and the pie chart compares a single years value for Italy. |
| **Collaboration** | Each member answered one question and created one visual based on there question. | This ensured fair workload for all members and enabled comprehensive analysis of Italy's environmental performance. |

## Italy's CO~2~ emissions vs its peers

```{r}
Data1 <- cO2_data |>
  filter(Year >= 1990, Entity %in% c("Italy", "France", "Germany", "United Kingdom", "European Union (27)", "China", "United States")) |> 
  rename(emissions_pc = "Annual CO₂ emissions (per capita)") |>
  mutate(Entity = fct_reorder2(Entity, Year, emissions_pc))

                             
ggplot(Data1, aes(x = Year, y = emissions_pc, color = Entity, group = Entity)) +
  geom_line(size = 1.2, alpha = 0.8) +
  
  geom_line(data = filter(Data1, Entity == "Italy"), 
            size = 1.4) +
  geom_point(data = filter(Data1, Entity == "Italy"),
             size = 2.2) +
  
  scale_color_brewer(palette = "Dark2") +

  labs(
    title = "Comparing Italy's annual CO2 Emissions per Capita (1990 - 2024)",
    y = "per Capita CO2 Emissions (Million Tonnes)",
    X = "",
    color = "Country"
  ) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    legend.position = "right"
  )
```

-   Italy consistently emits less CO₂ per capita than Germany and the wider European Union average, while remaining higher than France.
-   The United States emits nearly three times Italy’s CO2 per capita, but China follows an inverse trajectory, with rapid industrialization reaching almost 9 tonnes today.

::: footer
Data source: Global Carbon Budget (2025); Population based on various sources (2024)
:::

## Italy's Per capita CO₂ emissions by sector

```{r}

  
co2_sector <- clean_names(co2_sector, case = "snake") 
clCo2 <- co2_sector |> select(-per_capita_fugitive_emissions_of_carbon_dioxide_from_energy_production)



CS <- clCo2  |> rename(
  "Buildings" = per_capita_carbon_dioxide_emissions_from_buildings,
  "electricity and heat" = per_capita_carbon_dioxide_emissions_from_electricity_and_heat,
  "industry" = per_capita_carbon_dioxide_emissions_from_industry,
  "bunker fuels" = per_capita_carbon_dioxide_emissions_from_bunker_fuels,
  "land use change and forestry" = per_capita_carbon_dioxide_emissions_from_land_use_change_and_forestry,
  "emissions from transport" = per_capita_carbon_dioxide_emissions_from_transport,
  "manufacturing and construction" = per_capita_carbon_dioxide_emissions_from_manufacturing_and_construction,
  "other fuel combustion" = per_capita_carbon_dioxide_emissions_from_other_fuel_combustion)

CS <- CS |> pivot_longer(cols = c("Buildings",
                                   "electricity and heat",
                                   "industry",
                                   "bunker fuels",
                                   "land use change and forestry",
                                   "emissions from transport",
                                   "manufacturing and construction",
                                   "other fuel combustion"),
                          names_to = "sector",
                          values_to = "emission") |> 
  mutate(emission= round_half_up(emission, -1)) |> 
  arrange(desc(emission))

CS3 <- CS |> filter(entity == "Italy")

```

```{r}
latest <- CS3 |> filter(year == max(year)) |> arrange(desc(emission))
CS3$sector <- factor(CS3$sector, levels = latest$sector)

CS3 |> 
  ggplot(aes(x = year, y = emission, fill = sector)) +
  geom_area(alpha = 0.75) +
  labs(
    title = "Italy's per capita CO2 emissions by sector",
    subtitle = "1990–2022 • Stacked contributions of major sectors",
    y = "Emissions (kg per capita)",
    x = "",
    fill = "Sector"
  ) +
  scale_fill_paletteer_d("MexBrewer::Huida") +
  scale_x_continuous(breaks = seq(1990, 2022, 5)) +
  theme_minimal(base_size = 10) +
  theme(plot.background = element_rect(fill = "white"))+
  theme(
    legend.position = "right",
    legend.title = element_text(size = 9, face = "bold"),
    legend.text  = element_text(size = 8,face = "bold"),
    plot.title = element_text(size = 16, face = "bold", colour = "chocolate4"),
    plot.subtitle = element_text(size = 11),
    axis.title.y = element_text(size = 10),
    axis.text.x = element_text(size = 8, face = "bold"),
    axis.text.y = element_text(size = 8, face = "bold")
  )

```

-   Emissions peaked in mid-2000s, led by electricity/heat and transport.

-   Significant decline since then across nearly all sectors.

-   Trend reflects cleaner energy, efficiency gains, and economic shifts.

::: footer{.smaller}
Data source: Climate Watch (2025); Population based on various sources (2024)
:::

## Electricity and heat production for Italy vs its peers

```{r}
library(dplyr)
library(ggplot2)
elec_heat <- co2_sector |> 
  rename(country = entity, co2_from_heat_electricity = per_capita_carbon_dioxide_emissions_from_electricity_and_heat) |>
  filter(country %in% c("Italy", "European Union (27)", "France", "Germany", "China", "United States", "United Kingdom")) |>
  select(country,code,year,co2_from_heat_electricity)

max_elec_heat <- elec_heat |>
  filter(year == max(year))|>
  arrange(desc(co2_from_heat_electricity))

elec_heat$country <- factor(elec_heat$country, levels = max_elec_heat$country)

ggplot(elec_heat, aes(x = year, y = co2_from_heat_electricity, color = country))+
  geom_line(key_glyph = "timeseries",size=1.3, alpha = .5)+
  labs(title = "Per capita CO2 emissions from electricity and heat",
       subtitle = "Italy compared with UK, France, Germany, China and US", 
       x = "", 
       y= "Per capita CO2 emissions",
       colour = "Country")+
  theme_classic()+
  theme(plot.title = element_text(face = "bold"),
        axis.title = element_text(face = "bold"))+
  geom_line(key_glyph = "timeseries",
            data = elec_heat |> filter(country=="Italy"),
            aes(x = year, y = co2_from_heat_electricity),
            size = 2)
```

::: footer{.smaller}
Data source: Climate Watch (2025); Population based on various sources (2024)
:::

## what percentage of Italy's energy production come form renewable sources?

```{r}
# Cleaned/filtered data  
engery_clean <- co2_vs_renw |> 
  clean_names() |> 
  select(
    entity,
    year,
    annual_co2_emissions_per_capita,
    renewables_percent_electricity
  ) |> 
  filter(
    entity %in% c("Italy","United Kingdom", "France","China", "United States", "Germany"),
    year >= 1990 & year <= 2022 
  ) |> 
  filter(!is.na(renewables_percent_electricity))

```

```{r}
# Plotting  
ggplot(engery_clean, aes(
  x = year ,
  y = renewables_percent_electricity,
  colour = entity
)) +
  geom_line( 
    size = 1) +
  
  geom_point(size = 1.5, alpha = 1) +
  
  geom_line(
    data = engery_clean |> filter(entity == "Italy"),
    aes(x = year,
        y = renewables_percent_electricity),
    linewidth = 1.5,
  ) +
  geom_point(
    data = engery_clean |> filter(entity == "Italy"),
    aes(x = year,
        y = renewables_percent_electricity),
    size = 2.2
  ) +
  labs(
    title = "Renewable Electricity  (1990-2022)",
    x = "Year",
    y = "Renewable Share of Electricity",
    subtitle = "Italy compared with UK, France, Germany, China and US",
    colour = "Country",
  ) +
  scale_color_brewer(palette = "Dark2") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
    legend.title = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10)
)
```

-   Italy's renewable electricity share has seen a steady growth from 1990 to 2022, with massive growth starting from 2008.

-   Italy outperforms the US, France and China, while the UK and Germany show faster renewable growth peaks.

-   Italy's progress shows commitment towards cleaner electricity usage and their transition to achieve that goal.

::: footer{.smaller}
Data source: Global Carbon Budget (2025) and other sources
:::

## Percentage of Italy's energy production 

```{r}
electricty_pd <-  clean_names(electricty_pd)

clean_electricity_pd <- electricty_pd |> 
  rename(
    Other_Renewables = "other_renewables_excluding_bioenergy_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Bioenergry = "electricity_from_bioenergy_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Solar = "electricity_from_solar_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Wind = "electricity_from_wind_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Hydro = "electricity_from_hydro_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Nuclear = "electricity_from_nuclear_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Oil = "electricity_from_oil_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Gas = "electricity_from_gas_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked",
    Coal = "electricity_from_coal_t_wh_adapted_for_visualization_of_chart_electricity_prod_source_stacked")

italy_electricity_pd <- clean_electricity_pd |> 
    filter(entity == "Italy", year == 2022)

electricty_pd_longer <- italy_electricity_pd |> 
  pivot_longer(cols = c(Other_Renewables, Bioenergry, Solar, Wind, Hydro, Oil, Gas, Coal),
               names_to = "variables",
               values_to = "values") |> 
  select(-c(entity, code, year)) |> 
  arrange(values)

colors <- c("#BD3106FF", "#D9700EFF", "#E9A00EFF", "#EEBE04FF", "#5B7314FF", "#C3D6CEFF", "#89A6BBFF", "#454B87FF")

fig <- plot_ly(type='pie', labels=electricty_pd_longer$variables, values=electricty_pd_longer$values, 
               textinfo='percent',
               marker = list(colors = colors),
               insidetextorientation='radial',
               textfont = list(family = "Arial",
                               size = 15,
                               color = "white"))
fig <- fig |> add_annotations(
        y=1.06, 
        x=0.5, 
        text="Italy's Energy Production in 2022", 
        showarrow=F,
        font=list(size=15)
      )
fig <- fig |>
  layout(
    legend = list(
      font = list(size = 18),
      title = list(
        text = "Energy Types",
        font = list(size = 20)
      )
    )
  )
fig
```

## Slide 9: Sahadat

Table of Italy's Environmental Policy Comparison

| Comparative Region | Italy's Approach | Key Differences |
|------------------------|------------------------|------------------------|
| **(GE, FR, UK)** | Italy is the Circular Economy leader (recycling efficiency) but trails in renewable installation speed due to bureaucracy. | Resource Efficiency vs. Decarbonization Speed |
| **European Union** | Italy aligns with EU goals (PNIEC) but resists strict EV-only mandates, favouring "tech neutrality" (biofuels). | National Flexibility vs. Bloc Targets |
| **USA** | Italy relies on Carbon Pricing and regulation ("Sticks"); the USA relies on uncapped Subsidies/IRA ("Carrots"). | Regulation vs. Incentive |
| **China** | Italy has falling emissions but slow democratic permitting whereas China has rising emissions but rapid, state-mandated infrastructure build out. | Democratic Process vs. Command & Control. |

## Slide 10: Hermann

Conclusion
